---
title: "AVERAGE STATES ARIMAX MODEL - with log-back transformations"
author: "Victor Felix"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
  html_document: default
---

Here we run ensembles and single automatic ARIMAX models for forecasting weekly hospitalizations in 48 states on the contiguous U.S. These model use mean hospitalization across the U.S. states as exogenous variables. The function fits on a rolling window of previous 104 weeks for the state under analysis and a rolling windows with the same size with 1 week-lag for the exogenous variables to generate forecasts. These models include log-back transformations. It return some metrics that evaluate the performance of the models:target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles. The user can choose a  single best automatic ARIMAXs (auto=TRUE), or ensembles of 27 permutations of 0,1,2 pdq's (ES27=TRUE) or 64 permutations of 0,1,2,3 pdq's (ES64=TRUE). The user also chooses the number of weeks ahead for each forecast, and the size of the rolling window which is set as 2 years (104 weeks).

```{r}
knitr::opts_chunk$set(echo = TRUE)
```
!!!!!!!!!!!!!!!!!!!
LOADING THE DATASET
!!!!!!!!!!!!!!!!!!!
```{r}
library("tidyr")
library("MMWRweek")
library("data.table")
library("caret")
library("purrr")
library("dplyr")
library("tseries")
library("gtools")
library("forecast")
library("scoringutils")
library("covidHubUtils")
library("parallel")
library("future")#https://cran.r-project.org/web/packages/future/vignettes/future-4-issues.html
library("listenv")
library("epitools")
```
!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
LOADING DATASET AND FUNCTIONS 
!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

```{r}
###########################################
#       LOADING AND CLEANING THE DATASET  #        
###########################################

# Loads the AVERAGE states models
source("ES_AVERAGE_48.R", local = TRUE, chdir = TRUE)

# Loads the ILI dataset
my_data = read.csv("treated_influenza_hosp_dataframe_v2.csv")
my_data$target_end_date<-as.Date(my_data$target_end_date) # set the dates as dates

list_of_states <- split(my_data, my_data$state_name)

```

AUTO AVERAGE WEEK1 - ALL STATES

IF YOU WANT A FAST RESULT RUN ONLY THE AUTO_example, the others take longer.

```{r}

start_time <- Sys.time()

AUTO_AVERAGE_WEEK1_list <- mclapply(list_of_states, ES_AVERAGE, auto=TRUE, n_weeks_ahead=1,list_of_states=list_of_states,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# FINAL DATAFRAME
AUTO_AVERAGE_WEEK1 <- bind_rows(AUTO_AVERAGE_WEEK1_list, .id = "State")

```

AUTO AVERAGE WEEK2 - ALL STATES

```{r}
start_time <- Sys.time()

# RUN MODEL 
AUTO_AVERAGE_WEEK2_list <- mclapply(list_of_states, ES_AVERAGE, auto=TRUE, n_weeks_ahead=2,list_of_states=list_of_states,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)
# FINAL DATAFRAME
AUTO_AVERAGE_WEEK2 <- bind_rows(AUTO_AVERAGE_WEEK2_list, .id = "State")
```

AUTO AVERAGE WEEK3 - ALL STATES

```{r}
start_time <- Sys.time()

AUTO_AVERAGE_WEEK3_list <- mclapply(list_of_states, ES_AVERAGE, auto=TRUE, n_weeks_ahead=3,list_of_states=list_of_states,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)
# Combine the list of data frames into a single data frame with names as a column
AUTO_AVERAGE_WEEK3 <- bind_rows(AUTO_AVERAGE_WEEK3_list, .id = "State")
```

AUTO AVERAGE WEEK4 - ALL STATES

```{r}
start_time <- Sys.time()

AUTO_AVERAGE_WEEK4_list <- mclapply(list_of_states, ES_AVERAGE, auto=TRUE, n_weeks_ahead=4,list_of_states=list_of_states,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)
# Combine the list of data frames into a single data frame with names as a column
AUTO_AVERAGE_WEEK4 <- bind_rows(AUTO_AVERAGE_WEEK4_list, .id = "State")
```

```{r}
save.image("AVERAGE_MODELS_influenza_hospitalization.Rdata")
```

ES27 AVERAGE WEEK1 - ALL STATES

```{r}
start_time <- Sys.time()

ES27_AVERAGE_WEEK1_list <- mclapply(list_of_states, ES_AVERAGE, ES27=TRUE, n_weeks_ahead=1,list_of_states=list_of_states,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)
# FINAL DATAFRAME
ES27_AVERAGE_WEEK1 <- bind_rows(ES27_AVERAGE_WEEK1_list, .id = "State")
```

ES27 AVERAGE WEEK2 - ALL STATES

```{r}
start_time <- Sys.time()

ES27_AVERAGE_WEEK2_list <- mclapply(list_of_states, ES_AVERAGE, ES27=TRUE, n_weeks_ahead=2,list_of_states=list_of_states,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)
# FINAL DATAFRAME
ES27_AVERAGE_WEEK2 <- bind_rows(ES27_AVERAGE_WEEK2_list, .id = "State")
```

ES27 AVERAGE WEEK3 - ALL STATES

```{r}
start_time <- Sys.time()

ES27_AVERAGE_WEEK3_list <- mclapply(list_of_states, ES_AVERAGE, ES27=TRUE, n_weeks_ahead=3,list_of_states=list_of_states,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)
# FINAL DATAFRAME
ES27_AVERAGE_WEEK3 <- bind_rows(ES27_AVERAGE_WEEK3_list, .id = "State")
```

ES27 AVERAGE WEEK4 - ALL STATES

```{r}
start_time <- Sys.time()

ES27_AVERAGE_WEEK4_list <- mclapply(list_of_states, ES_AVERAGE, ES27=TRUE, n_weeks_ahead=4,list_of_states=list_of_states,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)
# FINAL DATAFRAME
ES27_AVERAGE_WEEK4 <- bind_rows(ES27_AVERAGE_WEEK4_list, .id = "State")
```

```{r}
save.image("AVERAGE_MODELS_influenza_hospitalization.Rdata")
```

ES64 AVERAGE WEEK1 - ALL STATES

```{r}
start_time <- Sys.time()

ES64_AVERAGE_WEEK1_list <- mclapply(list_of_states, ES_AVERAGE, ES64=TRUE, n_weeks_ahead=1,list_of_states=list_of_states, mc.cores=2,week_lag=1)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)
# FINAL DATAFRAME
ES64_AVERAGE_WEEK1 <- bind_rows(ES64_AVERAGE_WEEK1_list, .id = "State")
```

ES64 AVERAGE WEEK2 - ALL STATES

```{r}
start_time <- Sys.time()

ES64_AVERAGE_WEEK2_list <- mclapply(list_of_states, ES_AVERAGE, ES64=TRUE, n_weeks_ahead=2,list_of_states=list_of_states, mc.cores=2,week_lag=1)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)
# Combine the list of data frames into a single data frame with names as a column
ES64_AVERAGE_WEEK2 <- bind_rows(ES64_AVERAGE_WEEK2_list, .id = "State")

```

ES64 AVERAGE WEEK3 - ALL STATES

```{r}
start_time <- Sys.time()

ES64_AVERAGE_WEEK3_list <- mclapply(list_of_states, ES_AVERAGE, ES64=TRUE, n_weeks_ahead=3,list_of_states=list_of_states, mc.cores=2,week_lag=1)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)
# Combine the list of data frames into a single data frame with names as a column
ES64_AVERAGE_WEEK3 <- bind_rows(ES64_AVERAGE_WEEK3_list, .id = "State")

```

ES64 AVERAGE WEEK4 - ALL STATES

```{r}
start_time <- Sys.time()

ES64_AVERAGE_WEEK4_list <- mclapply(list_of_states, ES_AVERAGE, ES64=TRUE, n_weeks_ahead=4,list_of_states=list_of_states, mc.cores=2,week_lag=1)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)
# Combine the list of data frames into a single data frame with names as a column
ES64_AVERAGE_WEEK4 <- bind_rows(ES64_AVERAGE_WEEK4_list, .id = "State")

```

```{r}
save.image("AVERAGE_MODELS_influenza_hospitalization.Rdata")
```
